**Machine learning identifies unique taxa differentiating proximal and distal

### 2017-06-07 submission/manuscript.Rmd
discussion updated

how does bowel prep fit into all of this- do these match publisehd findings? sort of hard to differentiate bc so many humans?

@@

By revealing specific differences in microbial populations at each location in the gut via sampling an unprepared bowel, we can begin to form hypothesies about how specific host-microbe interactions can affect oncogenesis of proximal and distal CRC tumors. To this point, 16S community profiling studies are a good start but not enough information to probe these questions. Future studies examining metagenomic, metabolomic and host interactions at each site will be useful in further characterizing the host-microbe interactions in the development of proximal and distal colorectal cancer.

>>>>

Specific comparisons of our findings to previously published gut biogeography studies are additionally confounded by the use of bowel preparation methods in most other studies. A rare report of a matched-colonoscopy study that sampled 18 patient's colonic mucosa and lumenal contents prior to and after bowel cleansing (cite). This group found that mucosa and lumenal samples were distinguishable prior to bowel cleansing, but that bowel preparation resulted in an increase in shared OTUs between each site (Shobar 2016). Bowel cleansing not only made the samples harder to distinguish, it resulted in decreases in diversity across sites. Further, the differences were not great enough to overcome interpersonal differences between subjects. So, bowel preparation clearly induces bias into the microbes recovered from sampling the lumen or mucosa of a prepared bowel. Thus our findings of specific bacteria at each site of the colon are strengthened by the lack of bowel preparation.

@@

By revealing specific differences in microbial populations at each location in the gut via sampling an unprepared bowel, we can begin to form hypothesies about how specific host-microbe interactions can affect oncogenesis of proximal and distal CRC tumors. To this point, 16S community profiling studies are a good start but not enough information to probe these questions. Our sample set of matched proximal, distal, lumenal and mucosal samples from colons that have not undergone bowel preparation presents a unique opportunity to explore further Future studies examining metagenomic, metabolomic and host interactions at each site will be useful in further characterizing the host-microbe interactions in the development of proximal and distal colorectal cancer.
